This is an example of article title

### 2023-07-01 paper/paper.tex
chore: Additions to paper

  $$
  \begin{matrix}
    L &\triangleq& [m, h) \\
    R &\triangleq& [h, M)
  \end{matrix}
  $$

  \begin{figure}
    \centerline{
      \includegraphics[width=10pc]{figures/spatial-division.png}
      \includegraphics[width=10pc]{figures/spatial-tree.png}
    }
    \caption{ (a) Spatial division, (b) Spatial tree }
    \label{spatial-tree}
  \end{figure}

\end{definition}

@@

  % TODO Insert gaphs
\end{example}

@@

\subsection{Curse of high dimensionality}
Data of high dimensionality exponatialy get sparser as the number of dimensions grow.
The sparser the data the lower the probabilities are for two points to occupy the same
bin. In actuallity after the first recursion of the algorithm each element is most
probably assigned a unique bin, which defeats the purpose of the subdivision.

@@

\section{Experiments}

\subsection{Uniform data 2 dimensions}
\subsection{Uniform data 10 dimensions}
\subsection{MNIST data 10 dimensions}
\subsection{PBMC data 10 dimensions}
\subsection{PBMC 36 dimensions}

>>>>

  $$
  \begin{matrix}
    L &\triangleq& [m, h) \\
    R &\triangleq& [h, M)
  \end{matrix}
  $$
\end{definition}

@@

  \begin{figure}
    \centerline{
      \includegraphics[width=10pc]{figures/spatial-division.png}
      \includegraphics[width=10pc]{figures/spatial-tree.png}
    }
    \caption{ (a) Spatial division, (b) Spatial tree }
    \label{spatial-tree}
  \end{figure}
\end{example}

@@

\subsection{Curse of high dimensionality}
Data of high dimensionality exponatialy get sparser as the number of dimensions grow.
The sparser the data the lower the probabilities are for any two points occupying the same
bin.
\\\\
Assuming uniformly distributed $d$-dimensional $n$ points and a target tree of
$l$ levels, the expected number of points per leaf node is $E[n_\textrm{leaf}]$.

$$E[n_\textrm{leaf}] = \frac{n}{2^{l\cdot d}}$$

The expected number of points per leaf node drops exponentialy with $d$. This extends
to most distributions of points since the probability of a point belonging to a node is a
function of the $d$-dimensional area that the node occupies $A^{(d)}_l $.
In the case of uniform distribution the probability is directly proportional
to this area.

$$A^{(d)}_l = 2^{-l \cdot d}$$

In general the area is $A^{(d)}_l = s\cdot 2^{-l \cdot d}$ where $s$ is the scaling factor
of the original space. But since scaling does not affect the relative density of the
points the probability is not affected.

@@

\section{Experiments}
\subsection{Dimensionality scaling}
For $n=10^5$ uniformly distributed points we computed the average number of points in a
leaf node of the tree. The graph \ref{b_vs_d} shows that for increasingly higher
dimensional spaces the data get exponatialy sparser.

\begin{figure}[!ht]
  \centerline{
    \includegraphics[width=20pc]{figures/experiments/b_vs_d.png}
  }
  \caption{ Exponential decrease of mean number of points per node }
  \label{b_vs_d}
\end{figure}

\subsection{Uniform data 2 dimensions}

\begin{lstlisting}[language = Julia]
julia> d, n = size(X)
(2, 100000)

julia> l = 4;

julia> @time t = regural_bin(UInt128,
                             X,
                             l,
                             1;
                             dims=2);
0.025399 seconds (... 43.93% gc time)
\end{lstlisting}

For uniformly distributed data we expect that each leaf node will contain
$n_{\textrm{node}}$ points.
$$n_{\textrm{node}} = \frac{n}{2^{l\cdot d}}$$

In this case $n = 10^5$, $d = 2$ and $l = 4$. So the expected number of points within a
node is $n_{\textrm{node}} = \frac{10^5}{2^{4\cdot 2}} =390.625$.

\begin{figure}[!ht]
  \centerline{
    \includegraphics[width=20pc]{figures/uniform_2d/hist.png}
  }
  \caption{ Distribution of node length }
  \label{uniform-2d-hist}
\end{figure}

The curse of high dimensionality can be deduced by the formula of $n_{\textrm{node}}$.
The number of possible leaf nodes is $2^{l\cdot d}$ which grows exponentialy with $l$ and
$d$, meaning that for larger $d$ the total number of points $n$ must also grow exponentialy
to keep the ratio constant.

\subsection{Uniform data 10 dimensions}

\begin{lstlisting}[language = Julia]
julia> d, n = size(X)
(10, 100000)

julia> l = 1;

julia> @time t = regural_bin(UInt128,
                             X,
                             l,
                             1;
                             dims=2);
0.046737 seconds (2.48 k allocations: 30.789 MiB)
\end{lstlisting}

\subsection{MNIST data 10 dimensions}
\subsection{PBMC data 10 dimensions}
\subsection{PBMC 36 dimensions}
